---
title: "Estimating Value of Data Science Skills by Internet Scraping"
author: "Joel G. & Ben H."
date: "October 21, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction


## Purpose:

"Which are the most valued data science skills?""



## Analytic Strategy

1. Scrape job listings for data scientists from a jobs posting web site

2. Prepare, clean, and store data for analysis

3. Develop a list of data science hard and soft skills

4. Analyze:

    - Find the most commonly listed skills
    
    - Where salary information is available, compute the mean salary for each skill
    
    
    
# Preparation

## Data Collection

* Source: Indeed.com

* 'data scientist' job postings across five cities:
    
    - New York City
    - San Francisco
    - Chicago
    - Washington, D.C.
    - Seattle
    
* Scraped ~500 listings
    - But due to duplication, final $n = 137$


## Scraping: Query Strings

* We chose to scrape Indeed.com, due to their easy and clean query strings

| element      | example                      |
|--------------|------------------------------|
| base URL     | https://www.indeed.com/jobs? |
| search query | q=data+scientist             |
| location     | l=New+York%2C+NY             |
| page         | start=10                     |



## Scraping: Concatenating Query Strings

Putting it altogether, e.g.:

`https://www.indeed.com/jobs?q=data+scientist&l=New+York%2C+NY&start=10`

returns the second page of job listings for `data scientist` located in New York.



## Scraping Algorithm

1. For each `city` in the list of `cities`:
    1. Get some page of results, ` pages = seq(0, 100, 10) = (0, 10, 20, ..., 100)`.
    2. For each `page` in `pages`:
        1. Assemble a URL of job results
        2. Read the HTML of the resulting page: `rvest::read_html(results_url)`
        3. Extract the approximately 15 links for job postings contained in the HTML: **`extract_listing_urls(results_page_html)`**
        4. Use `sapply` to **`extract_listing_data()`** from each of these links
        5. Convert to data frame, save it as TSV
        
        
        
## Scraping Algorithm: Results

50 TSVs saved in the `./data/raw/` directory



## Data Cleaning

1. Remove NA values from the data frame.
2.  Remove duplicate job entries. 
3. Extract selected soft skills and hard skills obtained from reviewing literature.
4. Create a column for each soft and hard skill. When one of the keywords appeared on a job posting a TRUE value will be reflected in the corresponding column and row. 
4. Using the gather function to convert data from wide to long.
5. Convert True to 1 and False to 0 to count True values. 
6. Group the soft and hard skills and count how many times each soft and hard skill was found in each of the job postings. 





## Data Cleaning: Salary Information

* Salary data from Indeed comes in various formats
* `extract_salary()` uses `stringr` and Regular expressions to extract this information, depending on if it's a range of salaries or a single salary

```{r, eval=FALSE, echo=TRUE}
if (contains_range == TRUE) {
    text_range <- unlist(str_extract_all(s, '\\$[0-9,]+'))
    numeric_range <- as.numeric(str_remove_all(text_range, ',|\\$'))
    return(mean(c(numeric_range[1], numeric_range[2])))
} else {
    as_text <- str_extract_all(s, '\\$[0-9,]+')
    as_numeric <- as.numeric(str_remove_all(as_text, ',|\\$'))
    return(as_numeric)
}
```




## Data Storage

A SQL file was created with the table names.  The date scraped field did not load initially so the date field needed to be formatted to '%m/%d/%Y'. The soft skills and hard skills were also stored in the SQL database. 


# Analysis



## Soft and Hard Skills

* Assembled from reviewing literature on the Internet
* Soft skills:
    - Attention to detail
    - Communication
    - Interpersonal abilities
    - Leadership
    - Problem solving
    - Time management
* Hard skills included technical skills, like R, Python, data engineering, Java, etc.



## Analysis: By Count

* Soft skills are the character traits or interpersonal skills that affect your ability to work and in interact with others.  Ten (10) soft skills were selected from the following website - https://resumegenius.com . A search was a conducted through Data Scientist job postings that were advertised on Indeed.com. The soft skills:  communication, problem solving, time management, leadership, and attention to detail are found more frequently on job postings advertised on Indeed.com. Based on the number of occurrences of the foregoing soft skills, it can be assumed that these soft skills are most sought after by employers of data scientist positions. 

* According to resume genius.com, hard skills are the specific knowledge and abilities that are learned through education and training.  Similarly to the soft skills, twenty three (23) hard skills were selected from Indeed.com . The hard skills python, spark, R, hadoop, SQL, Big Data, Java are found more frequently on job postings advertised on Indeed.com. These hard skills it can also be assumed that these hard skills are most sought after by employers of data scientist positions. 




## Analysis: By Salary

* Only 22 of the original 134 job listings contained salary information, or about 16 percent

* Job listings were made 'long' in the form of `(job_id, skill, salary)`

* Then: Use `dplyr::summarize()` to find the average salary for each skill



## Analysis: Soft Skill Salaries

| rank | skill               | avg. salary ($) |
|-----:|---------------------|----------------:|
|    1 | time management     |         120,500 |
|    2 | communication       |         118,278 |
|    3 | problem solving     |         110,700 |
|    4 | attention to detail |         109,500 |
|    5 | interpersonal       |         106,000 |
|    6 | leadership          |          90,000 |



## Analysis: Hard Skill Salaries (Top 5)

| rank | skill                      | avg. salary ($) |
|-----:|----------------------------|----------------:|
|    1 | tensorflow (deep learning) |         132,250 |
|    2 | SQL                        |         122,281 |
|    3 | data mining                |         122,000 |
|    4 | Hive                       |         121,300 |
|    5 | Python                     |         120,417 |
|  ... | ...                        |             ... |



## Analysis: Hard Skill Salaries (Bottom 5)

| rank | skill       | avg. salary ($) |
|-----:|-------------|----------------:|
|  ... | ...         |             ... |
|   14 | Pig         |         110,000 |
|   15 | Java        |         105,375 |
|   16 | R           |         101,700 |
|   17 | Shell/bash  |          93,750 |
|   18 | pivot table |          57,500 |



## Analysis: Soft v. Hard Skills

* We calculated the average value of soft skills altogether and hard skills algother
    - Average hard skill salary: $116,432
    - Average soft skill salary: $112,328
    
* Given the small difference in average salaries and small sample size, we were unable to find evidence that either hard or soft skills are valued more by employers





# Conclusion

## Conclusions

[JOEL]

* Deep learning is the most highly valued hard skill

* Time management and communication are the most valued hard skills
    - Whereas 'leadership' is the least valued
    
* Python skill is valued by around $20,000 greater than R skill

* Accessing small and large data---SQL and Hive, respectively---remain an import skill for data scientists



## Further Work

* Larger sample size!

* With only 22 listings with salaries, that analysis especially cannot be considered definitive



## Challenges

* Scraper duplicated a lot of job postings, drastically decreasing our sample size

* Windows and Mac OS use different conventions for file paths

* Not all team members were able to use Git

* Difficulty of storing data with long strings of text in character delimited files



## Lessons

* Working off the same repository is critical

* URL query strings can make scraping loads easier!

* Rvest is a very straightforward package for scraping in R

* Use `file.path()` when specifying file locations to make code OS-independent

* When dealing with long strings of text, consider JSON for storage



## Team Responsibilities

* _Joel_: Data cleaning, data storage and SQL, analysis by count, some writing, graphics

* _Ben_: Scraping, analysis by salary, graphics, some writing


